---
title: "An Introduction to ggplot2 Part Two:  Legends"
author: "Jeff Atkins"
date: "October 24, 2017"
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

* For everything in this section, and for a lot of things moving forward, I have always found the [R Graphics Cookbook from Winston Chang](http://www.cookbook-r.com/Graphs/) to be invaluable. Most things are online, but buy it if ya can.
&nbsp;

&nbsp;


#### Begin by Initating ggplot2

```{r require-ggplot2, echo = FALSE, message = FALSE}
require(ggplot2)
```

## Legends
You have already succesfully made some pretty amazing plots using the `iris` dataset. Now it is time we really make them our own. One of the great things about ggplot2 (and R too) is not only can we express our data in a clear and concise manner, we can also personalize and customize those plots. Here is our plot from earlier:

```{r remembering-plot}
ggplot(iris, aes(x = Petal.Width, y = Petal.Length, color = Species))+
  geom_point(size = 4, alpha = 0.5)+
  xlab("Petal Width (cm)")+
  ylab("Petal Length (cm)")+
  theme_classic()
```

Not a fan of a legend to the side? No problem. If you use the very powerful `theme()` code, you can do some cool stuff. 

```{r legend-top}
ggplot(iris, aes(x = Petal.Width, y = Petal.Length, color = Species))+
  geom_point(size = 4, alpha = 0.5)+
  xlab("Petal Width (cm)")+
  ylab("Petal Length (cm)")+
  theme_classic()+
  theme(legend.position = "top")
```


```{r legend-bottom}
ggplot(iris, aes(x = Petal.Width, y = Petal.Length, color = Species))+
  geom_point(size = 4, alpha = 0.5)+
  xlab("Petal Width (cm)")+
  ylab("Petal Length (cm)")+
  theme_classic()+
  theme(legend.position = "bottom")
```


```{r inset1}
ggplot(iris, aes(x = Petal.Width, y = Petal.Length, color = Species))+
  geom_point(size = 4, alpha = 0.5)+
  xlab("Petal Width (cm)")+
  ylab("Petal Length (cm)")+
  theme_classic()+
  theme(legend.justification=c(0,0), legend.position=c(0,0)) # Put bottom-left corner of legend box in bottom-left corner of graph
```


```{r inset2}
ggplot(iris, aes(x = Petal.Width, y = Petal.Length, color = Species))+
  geom_point(size = 4, alpha = 0.5)+
  xlab("Petal Width (cm)")+
  ylab("Petal Length (cm)")+
  theme_classic()+
 theme(legend.justification=c(1,0), legend.position=c(1,0)) # Put bottom-right corner of legend box in bottom-right corner of graph
```
&nbsp;

&nbsp;


But wait, there's more! We can change around what our legend looks like and even change the color scheme of our data

```{r fun-time}
ggplot(iris, aes(x = Petal.Width, y = Petal.Length, color = Species))+
  geom_point(size = 4, alpha = 0.5)+
  xlab("Petal Width (cm)")+
  ylab("Petal Length (cm)")+
  theme_classic()+
  scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9"))+
 theme(legend.justification=c(1,0), 
       legend.position=c(1,0),
       legend.background = element_rect(fill="grey"))
```
                   
                   
```{r fun-time2}
ggplot(iris, aes(x = Petal.Width, y = Petal.Length, color = Species))+
  geom_point(size = 4, alpha = 0.5)+
  xlab("Petal Width (cm)")+
  ylab("Petal Length (cm)")+
  theme_classic()+
  scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9"), 
                       name="Species",
                       breaks=c("setosa", "versicolor", "virginica"),
                       labels=c("setosa", "versicolor", "virginica"))+
 theme(legend.justification=c(1,0), 
       legend.position=c(1,0))
```      
&nbsp;

&nbsp;

&nbsp;

&nbsp;


#### Challenge!

With all of this new found knowledge, using the `iris` dataset, I want you to produce the ugliest plot you can. 
